SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-12 docs/questioning.tex
some corrections

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps-new.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there are
corresponding transition numbers for
every attack. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub} (at the user's browser):~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc s}t{\sc b} (en route from the server to the browser):~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
Point {\sc e}i{\sc q} (embedded in a query en route to the database):~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point i{\sc db} (in the database):~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

A number of the frameworks define policies
based on information and features provided by
the Document Object Model ({\sc dom}) of a web page.
Specifically, developers must place all legitimate
scripts inside {\sc html} elements like {\tt div}.
The web browser (Point {\sc ub}), parses the
{\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts
are treated according to the policies defined
on the server. Frameworks that support this
functionality are {\sc beep}~\cite{TNH07}
and {\sc dsi}~\cite{NSS06}. The main problem with
the above mechanisms is that they are vulnerable
to non-persistent {\sc xss} attacks that alter
the {\sc dom} of an already rendered
page~\cite{APKLM10}.
In addition, a recent variation of JavaScript
injection attacks known as return-to-JavaScript
attack can be used to bypass the above mechanisms
as described extensively in reference~\cite{APKLM10}.
% Such attacks share
% many similarities with the {\it return-to-lib}
% attack~\cite{SPPGMMB04}. 
%% PL Explain what it does, I guess it's what you go on to say
%% directly below?
% When performing an attack like this the malicious user transfers the
% execution to points that contain script code which in turn is used to
% perform the attack.

@@

Policy enforcement mechanisms that detect {\sc csrf}
attacks are usually implemented in the form of a
server-side proxy (Point {\sc s}t{\sc b})
interposing on the client-server communication, and modifying
it. {\it NoForge}~\cite{JKK06a}
parses the {\sc html} server responses
and adds a token to every {\sc url} referring to this
server. Then it associates the token with the cookie
representing the session {\sc id} for the application.
When a request is received, the mechanism checks
if the request contains the token related
to the session {\sc id}. A disadvantage of NoForge
is that {\sc html} that is dynamically created within
the browser will not include the token.
Thus, sites that create their {\sc html} on the client
will remain vulnerable. In addition, it does not
support cross-origin requests.
{\it j{\sc csrf}}~\cite{PS11}
has a very similar functionality and meanwhile
addresses the above problems.
Finally, {\it CsFire}~\cite{DDHPJ10}
examines cross-domain interactions, to design a
cross-domain policy on the client's side (Point {\sc ub}).
The policy is based on the concept of a relaxed
same-origin policy that allows communication between
sub-domains of the same registered domain.

@@

In the case of {\sc xss} attacks that are based on JavaScript or {\sc
  html}, consider a {\sc xor} function that encodes all source of a
web page on the server-side and then, on the client-side, the web
browser decodes the source by applying the same function again (Point
{\sc ub}). Variations of this approach include {\it
  Noncespaces}~\cite{GC09} and {\it x{\sc js}}~\cite{APKLM10}, which
randomize the instruction set of {\sc html} and JavaScript
respectively. Contrary to x{\sc js}, in Noncespaces administrators
must set specific policies in a manner similar to a firewall
configuration language. This fact, together with the insufficient
testing of the mechanism (see Table~\ref{tab:comp}) brings up
questions concerning its accuracy. {\it {\sc
    sm}ask}~\cite{JB07} is another framework that was inspired by {\sc
  isr}. To detect {\sc xss} attacks it searches for {\sc html} and
JavaScript keywords within the application's legitimate code. This is
done before the processing of an {\sc http} request. When it finds one
it adds a token to it. This results to a ``code mask". Then, before
sending the resulting {\sc html} data to the web user, the framework
searches the data for illegal code by using the same keywords (Point
{\sc s}t{\sc b}). Since all legitimate code has been ``masked", the injected
code can be identified. The pre- and post-processing of the code
though, may add a significant overhead to the application.
Unfortunately, the authors of {\sc sm}ask did not provide measurements
regarding the computational performance of the tool (see
Table~\ref{tab:comp}).

@@

  &   {\it Vogt et al.}~\cite{VFJKKV07} & 322 & ({\sc nq},{\sc nq},{\sc nq},{\sc na})\_r & {\sc nq} ({\sc s}) & {\sc xss} \\

@@

However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript
attack~\cite{APKLM10}.

@@

{\sc xss-guard}~\cite{BV08} is a training scheme that also employs
parse-tree validation. During training, the scheme maps legitimate
scripts to {\sc http} responses. During production {\sc xss-guard}
retrieves for every script included in a response its parsed tree and
checks if it is one of those previously mapped to this response. Apart
from the comparison of the parsed trees, {\sc xss-guard} checks also
for an exact match of lexical entities. To achieve this, the scheme
utilizes that data structures of Firefox's JavaScript engine (Point
{\sc ub}). However, string literals are not compared literally and
therefore a false negative result could occur if an adversary performs
a mimicry attack~\cite{WS02}. For example, consider a banner rotator
that every time it runs, it creates a value that depends on the
current date and the length of the array that contains the references
of the various images to be displayed. Then, based on this value it
shows a specific image to a user. In a vulnerable web site that allows
users to post data and contains this banner rotator, a malicious user
could create and store a script that has the same code structure, with
the same JavaScript keywords contained in the rotator script.
%% PL You mean like the rotator?
In this script the attacker could also include references to tiny
images hosted on a web server that is maintained by him in order to
retrieve the {\sc ip} addresses of the users that visit the vulnerable
site.

{\it Blueprint}~\cite{LV09} is a policy enforcement framework that
also utilizes parsed trees to detect {\sc xss} attacks. To guarantee
that untrusted content is not executed, Blueprint generates on the
server-side a parsed tree from untrusted {\sc html} to ensure that it
does not contain any dynamic content. Then the parsed tree is
transfered to the document generator of the browser (Point {\sc ub})
where untrusted browser parsing behavior is ruled out. Blueprint is an
efficient countermeasure but it imposes a significant overhead due to
its extensive parsing (see Table~\ref{tab:comp}).

@@

In frameworks that deal with JavaScript-driven {\sc xss} attacks we
see that attack detection may take place either on the server-side or
on the client-side. In the first case, a proxy is placed on the server
to examine the server's responses before they reach the user's web
browser. In the second case a modified browser or a library that is
securely downloaded from the server checks the responses for potential
attacks. We find that the tendency so far is to create frameworks that
perform detection on the client side: in particular, 11 out of 19
frameworks (57,8\%) detect JavaScript-driven {\sc xss} attacks on the
client's side. Note that the majority of the mechanisms that detect
such attacks on the server-side can also detect {\sc dsl} code
injection attacks (7 out of 8).

>>>>

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps-new.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there are
corresponding transition numbers for
every attack. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub} (at the user's browser):~\cite{KJKV09,LV09,TNH07,NSS06,APKLM10,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc s}t{\sc b} (en route from the server to the browser):~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08,PS11},
Point {\sc e}i{\sc q} (embedded in a query en route to the database):~\cite{BWS05,SW06,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point i{\sc db} (in the database):~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

A number of the frameworks define policies
based on information and features provided by
the Document Object Model ({\sc dom}) of a web page.
Specifically, developers must place all legitimate
scripts inside {\sc html} elements like {\tt div}.
The web browser (Point {\sc ub}), parses the
{\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts
are treated according to the policies defined
on the server. Frameworks that support this
functionality are {\sc beep}~\cite{TNH07}
and {\sc dsi}~\cite{NSS06}. The main problem with
these mechanisms is that they do
not examine the script's location inside the web
document. Based on this fact, attackers
can perform mimicry attacks~\cite{WS02}.
Specifically, they can execute legitimate scripts,
but not as intended by the original design of
the developers. This is extensively described
in reference~\cite{APKLM10}. In addition,
the authors of~\cite{APKLM10} describe
another recent variation of JavaScript
injection attacks that can be used to bypass the
above mechanisms, known as return-to-JavaScript
attacks.
% Such attacks share
% many similarities with the {\it return-to-lib}
% attack~\cite{SPPGMMB04}. 
%% PL Explain what it does, I guess it's what you go on to say
%% directly below?
% When performing an attack like this the malicious user transfers the
% execution to points that contain script code which in turn is used to
% perform the attack.

@@

Policy enforcement mechanisms that detect {\sc csrf}
attacks are usually implemented in the form of a
server-side proxy (Point {\sc s}t{\sc b})
interposing on the client-server communication, and modifying
it. {\it NoForge}~\cite{JKK06a}
parses the {\sc html} server responses
and adds a token to every {\sc url} referring to this
server. Then it associates the token with the cookie
representing the session {\sc id} for the application.
When a request is received, the mechanism checks
if the request contains the token related
to the session {\sc id}. A disadvantage of NoForge
is that {\sc html} that is dynamically created within
the browser will not include the token.
Thus, sites that create their {\sc html} on the client
will remain vulnerable. In addition, it does not
support cross-origin requests.
{\it j{\sc csrf}}~\cite{PS11} has a similar
functionality and meanwhile addresses the above problems.
Finally, {\it CsFire}~\cite{DDHPJ10}
examines cross-domain interactions, to design a
cross-domain policy on the client's side (Point {\sc ub}).
The policy is based on the concept of a relaxed
same-origin policy that allows communication between
sub-domains of the same registered domain.

@@

In the case of {\sc xss} attacks that are based on JavaScript or {\sc
  html}, consider a {\sc xor} function that encodes all source of a
web page on the server-side and then, on the client-side, the web
browser decodes the source by applying the same function again (Point
{\sc ub}). Variations of this approach include {\it
  Noncespaces}~\cite{GC09} and {\it x{\sc js}}~\cite{APKLM10}, which
randomize the instruction set of {\sc html} and JavaScript
respectively. Contrary to x{\sc js}, in Noncespaces administrators
must set specific policies in a manner similar to a firewall
configuration language. {\it {\sc sm}ask}~\cite{JB07} is another
framework that was inspired by {\sc isr}. To detect {\sc xss}
attacks it searches for {\sc html} and
JavaScript keywords within the application's legitimate code. This is
done before the processing of an {\sc http} request. When it finds one
it adds a token to it. This results to a ``code mask". Then, before
sending the resulting {\sc html} data to the web user, the framework
searches the data for illegal code by using the same keywords (Point
{\sc s}t{\sc b}). Since all legitimate code has been ``masked", the injected
code can be identified. The pre- and post-processing of the code
though, may add a significant overhead to the application.
Unfortunately, the authors of {\sc sm}ask did not provide measurements
regarding the computational performance of the tool (see
Table~\ref{tab:comp}).

@@

  &   {\it Vogt et al.}~\cite{VFJKKV07} & 322 & ({\sc nq},{\sc nq},{\sc nq},{\sc na})\_r & {\sc nq} ({\sc c}) & {\sc xss} \\

@@

However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect mimicry attacks 
where legitimate scripts can be executed by
attackers, but not in the way intended by
the the developers~\cite{APKLM10}.

@@

{\sc xss-guard}~\cite{BV08} is a training scheme that also employs
parse-tree validation. During training, the scheme maps legitimate
scripts to {\sc http} responses. During production {\sc xss-guard}
retrieves for every script included in a response its parsed tree and
checks if it is one of those previously mapped to this response. Apart
from the comparison of the parsed trees, {\sc xss-guard} checks also
for an exact match of lexical entities. To achieve this, the scheme
utilizes that data structures of Firefox's JavaScript engine (Point
{\sc ub}). However, string literals are not compared literally and
therefore a false negative result could occur if an adversary performs
the following attack: Consider a banner rotator
that every time it runs, it creates a value that depends on the
current date and the length of the array that contains the references
of the various images to be displayed. Then, based on this value it
shows a specific image to a user. In a vulnerable web site that allows
users to post data and contains this banner rotator, a malicious user
could create and store a script that has the same code structure, with
the same JavaScript keywords contained in the rotator script.
%% PL You mean like the rotator?
In this script the attacker could also include references to tiny
images hosted on a web server that is maintained by him in order to
retrieve the {\sc ip} addresses of the users that visit the vulnerable
site.

{\it Blueprint}~\cite{LV09} is a policy enforcement framework that
also utilizes parsed trees to detect {\sc xss} attacks. To guarantee
that untrusted content is not executed, Blueprint generates on the
server-side a parsed tree from untrusted {\sc html} to ensure that it
does not contain any dynamic content. Then the parsed tree is
transfered to the document generator of the browser (Point {\sc ub})
where untrusted browser parsing behavior is ruled out.
Blueprint is an efficient countermeasure but it imposes an
overhead due to its extensive parsing (see Table~\ref{tab:comp}).

@@

In frameworks that deal with {\sc xss} attacks we
see that attack detection may take place either on the server-side or
on the client-side. In the first case, a proxy is placed on the server
to examine the server's responses before they reach the user's web
browser. In the second case a modified browser or a library that is
securely downloaded from the server checks the responses for potential
attacks. We find that the tendency so far is to create frameworks that
perform detection on the client side: in particular, 14 out of 23
frameworks (60,8\%) detect {\sc xss} attacks on the
client's side. Note that the majority of the mechanisms that detect
such attacks on the server-side can also detect {\sc dsl} code
injection attacks (6 out of 9).
